---
title: "search-strategy"
author: "Rex Parsons"
date: "`r Sys.Date()`"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(naniar)
```


```{r}
load(file='data/all_studies.rda')
```

### list() in `primary_outcome_description` and `secondary_outcome_description` represent missing fields. 
They are the content below the outcome measures in the registration. Here, I clean up the character cols and remove these "list()" before assessing missingness. Then, since these descriptions appear to only be more verbose versions of the `[primary/secondary]_outcome_measures`, I concatenate them together and drop the old column.

```{r}
head(dat, 20) %>% 
  select(primary_outcome_measures, primary_outcome_description, 
         secondary_outcome_measures, secondary_outcome_description) %>%
  knitr::kable()

dat_cleaned <- dat %>%
  mutate(
    across( # remove all the 'list()'s from outcome descriptions
      c("secondary_outcome_description", "primary_outcome_description"),
      function(x) str_remove_all(x, "list\\(\\)")
    ),
    across(where(is.character), str_trim), # remove white space
    across(where(is.character), function(x) ifelse(nchar(x) == 0, NA, x) ), # replace empty character fields with NA,
    across(submitted:updated, function(x) as.Date.character(x, format="%B %d, %Y"))
  )

head(dat_cleaned, 20) %>% 
  select(primary_outcome_measures, primary_outcome_description, 
         secondary_outcome_measures, secondary_outcome_description) %>%
  knitr::kable()
```

# visualise the missingness

```{r}
n_sample <- 10000
df_sample <- dat_cleaned[sample(1:nrow(dat_cleaned), size=n_sample, replace=FALSE), ]
```

Lots of missingness in `last_known_status`. This variable is only ever complete when `overall_status == "Unknown status"`

```{r}
vis_miss(df_sample)
gg_miss_fct(x = select(df_sample, overall_status, last_known_status), fct = overall_status)
```

Other vars with relatively high amounts of missingness (ignoring `[primary/secondary]_outcome_description`) include:

* keywords
* detailed_summary
* study_design_purpose
* study_design_allocation

```{r}
df_sample2 <- select(df_sample, -last_known_status)
gg_miss_upset(df_sample2, nsets=10)
gg_miss_var(df_sample2, show_pct = TRUE)
```

# Experiment with some search-strategies

Concatenate outcome descriptions to measures, and drop the descriptions cols
```{r}
# n_sample <- 1000
# dat_search <- dat_cleaned[sample(1:nrow(dat_cleaned), size=n_sample, replace=FALSE), ] 
dat_search <- dat_cleaned # don't use sample sample later


dat_search2 <- dat_search %>%
  # concatenate primary and secondary outcome measures and descriptions
  mutate(
    across(
      c("primary_outcome_measures", "primary_outcome_description",
        "secondary_outcome_measures", "secondary_outcome_description"),
      ~replace_na(., "")
    ),
    primary_outcome = str_c(primary_outcome_measures, '; ', primary_outcome_description),
    secondary_outcome = str_c(secondary_outcome_measures, '; ', secondary_outcome_description)
  ) %>%
  select(
    -c("primary_outcome_measures", "primary_outcome_description",
       "secondary_outcome_measures", "secondary_outcome_description")
  ) %>%
  # concatenate official_title with brief_title, and brief_summary with detailed_summary
  mutate(
    across(
      c("official_title", "brief_title",
        "brief_summary", "detailed_summary"),
      ~replace_na(., "")
    ),
    title=str_c(official_title, '; ', brief_title),
    summary=str_c(brief_summary, '; ', detailed_summary)
  ) %>%
  select(
    -c("official_title", "brief_title",
       "brief_summary", "detailed_summary")
  )

```

# functions to apply search strategies
```{r}
to_lower_all <- function(data, apply=TRUE) {
  if(apply) {
    data %>% 
      mutate(across(everything(), tolower))
  } else {
    data
  }
}

fx_check_regex <- function(data, cols, pattern, indicator_colname, apply_to_lower=TRUE) {
  match_vec <- data %>%
    select(all_of(cols)) %>%
    to_lower_all(., apply=apply_to_lower) %>%
    mutate(across(
      cols, 
      function(x) str_detect(x, pattern=pattern)
      )) %>%
    rowSums > 0
  
  if(indicator_colname %in% names(data)) {
    data[[indicator_colname]] <- (data[[indicator_colname]] + match_vec) > 0
  } else {
    data[[indicator_colname]] <- match_vec > 0
  }
  data
}

# test example to show that piping and updating existing 'indicator_column' works
example_df <- data.frame(
  abstract = c("chest pain", "Chest pain", "angina", "prediction model"),
  title = c("test1", "test2", "chest pain in ...", "angina")
)

# example_df %>%
#   fx_check_regex(cols="abstract", pattern="chest pain", apply_to_lower = FALSE, indicator_colname = "heart_disease")
# 
# example_df %>%
#   fx_check_regex(cols="abstract", pattern="chest pain", indicator_colname = "heart_disease")
# 
# example_df %>%
#   fx_check_regex(cols=c("abstract", "title"), pattern="chest pain", indicator_colname = "heart_disease")
# 
# example_df %>%
#   fx_check_regex(cols=c("abstract", "title"), pattern="angina", indicator_colname = "heart_disease")

example_df %>%
  fx_check_regex(cols=c("abstract", "title"), pattern="angina", indicator_colname = "heart_disease") %>%
  fx_check_regex(cols=c("abstract", "title"), pattern="chest pain", indicator_colname = "heart_disease") %>%
  fx_check_regex(cols=c("abstract", "title"), pattern="prediction model", indicator_colname = "clinical_prediction_model") %>%
  mutate(include=heart_disease & clinical_prediction_model)

```

search cols:

* summary
* title
* primary_outcome
* secondary_outcome

target approx n=1000 included studies = 0.25%

```{r}
# n_sample <- 100000
# dat_search3 <- dat_search2[sample(1:nrow(dat_search2), size=n_sample, replace=F),]
dat_search3 <- dat_search2
# head(dat_search3)

all_cols <- c("summary", "title", "primary_outcome", "secondary_outcome")

dat_search_results <- 
  dat_search3 %>% 
  fx_check_regex(cols=all_cols, pattern="prediction model", indicator_colname = "cpm") %>% 
  fx_check_regex(cols=all_cols, pattern="risk score", indicator_colname = "cpm") %>% 
  fx_check_regex(cols=all_cols, pattern="prognostic model", indicator_colname = "cpm") %>%
  fx_check_regex(cols=all_cols, pattern="diagnostic model", indicator_colname = "cpm") %>%
  mutate(include=cpm) %>%
  filter(include==TRUE) 

dat_search_results %>%
  summarize(
    included_n = sum(include==TRUE),
    included_perc = scales::percent(included_n/nrow(dat_search3), accuracy=0.01)
  )

# write.csv(dat_search_results, "data/included_studies.csv", row.names=FALSE)
```

```{r}
preview_html <- function(code){
  tempDir <- tempfile()
  dir.create(tempDir)
  htmlFile <- file.path(tempDir, "index.html")
  writeLines(paste0('<html><body>',
                    code,
                    '</body></html>'), htmlFile)
  viewer <- getOption("viewer")
  viewer(htmlFile)
}
```

Search strings
```{r}

prediction_model_terms <- c(
  "machine learning","artificial intelligence","deep learning",
  "prediction model","predictive model","prediction score","predictive score",
  "warning score","risk score","risk prediction",
  "prognostic model","diagnostic model"
) %>% 
  str_c(collapse='|')

prognostic_terms <- "prognos*"
diagnostic_terms <- "diagnos*"

df_matches <- 
  dat_search3 %>% 
  fx_check_regex(cols=all_cols, pattern=prediction_model_terms, indicator_colname="cpm") %>%
  fx_check_regex(cols=all_cols, pattern=diagnostic_terms, indicator_colname="cpm_type_diagnostic") %>%
  fx_check_regex(cols=all_cols, pattern=prognostic_terms, indicator_colname="cpm_type_prognostic")


df_matches %>%
  filter(
    cpm,
    (study_type == "Observational" | study_type == "Observational [Patient Registry]"),
    (cpm_type_diagnostic | cpm_type_prognostic)
  ) %>% nrow()

df_interventional_matches <-
  df_matches %>%
  filter(
    cpm,
    !(study_type == "Observational" | study_type == "Observational [Patient Registry]"),
    (cpm_type_diagnostic | cpm_type_prognostic)
  ) %>% 
  mutate(summary = paste(
    sep="<br><br>", 
    paste0("PRIMARY OUTCOME:", primary_outcome),
    paste0("SECONDARY OUTCOME:", secondary_outcome), 
    paste0("SUMMARY:", summary)
  ),
  summary=str_replace_all(summary, '([[:upper:]]+?\\s?+[[:upper:]]+:)', '<br><br>\\1'),
  summary=str_replace_all(summary, '<br><br><br><br>', '<br><br>'),
  authors="", journal="clintrials") %>%
  mutate(issn="1234", volume="1234", issue="1234", pages="1234", day="1", month="2", year="2000", publisher="blah", pmc_id="1234", pubmed_id="1234", url="1234") %>%
  mutate(day=day(posted), month=month(posted), year=year(posted)) %>%
  select(key=id, title=title, authors, journal, abstract=summary, everything()) %>%
  mutate(across(where(is.character), function(x) str_replace_all(x, ",", ";"))) 



# formatting for rayyan
included_studies_highlight <- c(
  "machine learning","artificial intelligence","deep learning",
  "prediction model","predictive model","prediction score","predictive score",
  "warning score","risk score","risk prediction"
) %>% 
  str_c(collapse='|')

exclude_highlight <- c("robot[a-z]*", "biomarker[a-z]*")%>% 
  str_c(collapse='|')

diagnostic_highlight <- "diagnos[a-z]*"

prognostic_highlight <- "prognos[a-z]*"

include_term_style <- "background-color: #90EE90; color: black; font-weight:bold;"
diagnostic_term_style <- "background-color: yellow; color: black; font-weight:bold;"
prognostic_term_style <- "background-color: #E0B0FF; color: black; font-weight:bold;"
exclude_term_style <- "background-color: 	red; color: white; font-weight:bold;"



maybe_include <- c(3, 4, 5, 6, 7, 10)
df_interventional_matches %>% 
  slice(maybe_include) %>% 
  mutate(
    abstract=str_replace_all(
      abstract, 
      glue::glue('({included_studies_highlight})'), 
      glue::glue('<span style="{include_term_style}">\\1</span>')
    ),
    abstract=str_replace_all(
      abstract,
      glue::glue('({diagnostic_highlight})'),
      glue::glue('<span style="{diagnostic_term_style}">\\1</span>')
    ),
    abstract=str_replace_all(
      abstract,
      glue::glue('({prognostic_highlight})'),
      glue::glue('<span style="{prognostic_term_style}">\\1</span>')
    ),
    abstract=str_replace_all(
      abstract,
      glue::glue('({exclude_highlight})'),
      glue::glue('<span style="{exclude_term_style}">\\1</span>')
    )
  ) %>%
  pull(abstract) %>% 
  paste0(collapse = "<br><br><br><br>####################<br><br><br><br>") %>%
  preview_html()

```



```{r}

# df_matches %>% filter(cpm & (cpm_type_diagnostic | cpm_type_prognostic)) %>% View() 

df_matches <- 
  dat_search3 %>% 
  fx_check_regex(cols=all_cols, pattern=prediction_model_terms, indicator_colname="cpm") %>%
  fx_check_regex(cols=all_cols, pattern=diagnostic_terms, indicator_colname="cpm_type_diagnostic") %>%
  fx_check_regex(cols=all_cols, pattern=prognostic_terms, indicator_colname="cpm_type_prognostic") %>%
  fx_check_regex(cols=all_cols, pattern="develo[a-z]*", indicator_colname="study_type_development") %>%
  fx_check_regex(cols=all_cols, pattern="externa[a-z]* validat[a-z]*", indicator_colname="study_type_validation") %>%
  fx_check_regex(cols=all_cols, pattern="impact study", indicator_colname="study_type_impact")


df_matches2 <- 
  df_matches %>%
  select(id, starts_with("cpm_type")) %>%
  pivot_longer(starts_with("cpm_type"), names_transform = function(x) str_remove(x, "cpm_type_")) %>%
  filter(value) %>%
  group_by(id) %>%
  summarize(cpm_model_type = paste0(name, collapse="; ")) %>%
  left_join(select(df_matches, id, cpm), .)


df_matches3 <- 
  df_matches %>%
  select(id, starts_with("study_type_")) %>%
  pivot_longer(starts_with("study_type_"), names_transform = function(x) str_remove(x, "study_type_")) %>%
  filter(value) %>%
  group_by(id) %>%
  summarize(cpm_study_type = paste0(name, collapse="; ")) %>%
  left_join(select(df_matches2, id, cpm, cpm_model_type), .) 

# write.csv(df_matches3, "data/study_groupings.csv", row.names=F)

```


### Write data out in a way that can be imported to rayyan
```{r}

all_included_studies <- 
  df_matches %>%
  filter(
    cpm,
    (study_type == "Observational" | study_type == "Observational [Patient Registry]"),
    (cpm_type_diagnostic | cpm_type_prognostic)
  ) %>% 
  mutate(summary = paste(
    sep="<br><br>", 
    paste0("PRIMARY OUTCOME:", primary_outcome),
    paste0("SECONDARY OUTCOME:", secondary_outcome), 
    paste0("SUMMARY:", summary)
  ),
  summary=str_replace_all(summary, '([[:upper:]]+?\\s?+[[:upper:]]+:)', '<br><br>\\1'),
  summary=str_replace_all(summary, '<br><br><br><br>', '<br><br>'),
  authors="", journal="clintrials") %>%
  mutate(issn="1234", volume="1234", issue="1234", pages="1234", day="1", month="2", year="2000", publisher="blah", pmc_id="1234", pubmed_id="1234", url="1234") %>%
  mutate(day=day(posted), month=month(posted), year=year(posted)) %>%
  select(key=id, title=title, authors, journal, abstract=summary, everything()) %>%
  mutate(across(where(is.character), function(x) str_replace_all(x, ",", ";"))) 

# all_included_studies <- read.csv("data/rayyan/not used/included_studies.csv")
# write.csv(all_included_studies, "data/included_studies.csv", row.names=F) 
```

# test space for conditional text formatting
```{r}
all_included_studies %>% 
  slice(1:2) %>% 
  mutate(
    abstract=str_replace_all(
      abstract, 
      glue::glue('({included_studies_highlight})'), 
      glue::glue('<span style="{include_term_style}">\\1</span>')
    ),
    abstract=str_replace_all(
      abstract,
      glue::glue('({diagnostic_highlight})'),
      glue::glue('<span style="{diagnostic_term_style}">\\1</span>')
    ),
    abstract=str_replace_all(
      abstract,
      glue::glue('({prognostic_highlight})'),
      glue::glue('<span style="{prognostic_term_style}">\\1</span>')
    ),
    abstract=str_replace_all(
      abstract,
      glue::glue('({exclude_highlight})'),
      glue::glue('<span style="{exclude_term_style}">\\1</span>')
    )
  ) %>%
  pull(abstract) %>% 
  preview_html()


```


```{r}
all_included_studies_out <- 
  all_included_studies %>% 
  mutate(
    abstract=str_replace_all(
      abstract, 
      glue::glue('({included_studies_highlight})'), 
      glue::glue("<span style='{include_term_style}'>\\1</span>")
    ),
    abstract=str_replace_all(
      abstract,
      glue::glue('({diagnostic_highlight})'),
      glue::glue("<span style='{diagnostic_term_style}'>\\1</span>")
    ),
    abstract=str_replace_all(
      abstract,
      glue::glue('({prognostic_highlight})'),
      glue::glue("<span style='{prognostic_term_style}'>\\1</span>")
    ),
    abstract=str_replace_all(
      abstract,
      glue::glue('({exclude_highlight})'),
      glue::glue("<span style='{exclude_term_style}'>\\1</span>")
    )
  )

# write.csv(all_included_studies_out, "data/included_studies.csv", row.names=F) 
```

# link NCT to final decisions
```{r}

final_decisions <- readxl::read_xlsx("data/rayyan/final_screening_decisions.xlsx")
cbind(final_decisions[1:5,2])
final_decisions %>% head() %>% select(title)
all_included_studies_out %>% head() %>% select(title) %>% remove_rownames()

cbind(
  select(final_decisions, title_final=title, final_decision), 
  select(all_included_studies, title_included=title, key)
) %>%
  filter(title_final != title_included)

# order is maintained from output from rayyan so can simply rbind the NCT to the rayyan output:

final_decisions_with_NCT <- cbind(select(all_included_studies, NCT=key, title_old=title), final_decisions)
final_decisions_with_NCT %>% 
  select(title_old, title) %>%
  rbind(head(.), tail(.)) %>% 
  remove_rownames()

write.csv(final_decisions_with_NCT, file="data/rayyan/final_decisions_with_NCT.csv", row.names = FALSE)
write.csv(filter(final_decisions_with_NCT, final_decision=="Include"), file="data/rayyan/final_included_with_NCT.csv", row.names = FALSE)

final_decisions_with_NCT_only <-
  final_decisions_with_NCT %>%
  filter(final_decision=="Include") %>%
  select(NCT, final_decision)
  
  
write.csv(final_decisions_with_NCT_only, file="data/rayyan/final_included_NCT_only.csv", row.names = FALSE)
```



```{r}


# df_matches3 %>%
#   filter(cpm_model_type == "diagnostic") %>%
#   select(id) %>%
#   inner_join(., all_included_studies, by=c("id" = "key")) %>%
#   write.csv("data/included_diagnostic_studies.csv", row.names=F) 

# df_matches3 %>%
#   filter(cpm_model_type == "prognostic") %>%
#   select(id) %>%
#   inner_join(., all_included_studies, by=c("id" = "key")) %>%
#   write.csv("data/included_prognostic_studies.csv", row.names=F) 

# have to find-and-remove "'s in the csv file before uploading to Rayyan for it to work. Tried quote=F in write.csv() but that didn't work.

df_matches3 %>%
  filter(cpm) %>%
  group_by(cpm_model_type) %>%
  summarize(n=n())

```
## Sankey plot
note: might swap to geom_flow() if we think percentages would be useful 
https://stackoverflow.com/questions/64146971/sankey-alluvial-diagram-with-percentage-and-partial-fill-in-r
```{r}
library(ggsankey)
df_matches3 %>%
  filter(cpm) %>%
  mutate(across(everything(), function(x) ifelse(is.na(x), "UNKNOWN", x) )) %>%
  make_long(cpm_model_type,cpm_study_type) %>%
  ggplot(aes(
    x=x, next_x=next_x, node=node, next_node=next_node, 
    fill=factor(node),label = node
  )) +
  geom_sankey(flow.alpha = 0.6) +
  geom_sankey_label() +
  theme_sankey(base_size = 16) +
  theme(legend.position = "none") +
  labs(x="")
```

### Check the proportion that mention "development" vs "external validation" vs "impact study"
```{r}
dat_categories <- 
  dat_search_results %>%
  fx_check_regex(cols=all_cols, pattern="develo[a-z]*", indicator_colname = "development") %>% 
  fx_check_regex(cols=all_cols, pattern="externa[a-z]* validat[a-z]*", indicator_colname = "validation") %>% 
  fx_check_regex(cols=all_cols, pattern="impact study", indicator_colname = "impact") 

dat_categories %>%
  select(id, development:impact) %>%
  pivot_longer(!id) %>%
  filter(value) %>%
  group_by(id) %>%
  summarize(type=paste0(name, collapse="; ")) %>%
  left_join(select(dat_categories, id), .) %>%
  plyr::count("type")

```